Peer-to-Peer Permissionless Consensus via Authoring Reputation

### 2023-05-10 freechains/freechains.tex
...

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in a consistent order, and
(iii) are respectful and on topic.
In a centralized system, items (i) and (ii) are trivially achieved assuming
availability and delivery order in the service, while for item (iii), users
have to trust the service to moderate content.
In a decentralized setting, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively replicate and
disseminate posts among peers until they reach all
users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to eradicate Sybil attacks~\cite{p2p.sybil}, which are the
major threat to decentralized applications in general: without consensus, it
is not possible, \emph{at the protocol level}, to distinguish between correct
and malicious users in order to satisfy item (iii).
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illegal content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybils through consensus.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects; and
    (iii) they impose an external economic cost to use the protocol.
These issues threaten our original decentralization goals.
In particular, a unique timeline implies that all Internet content should be
subject to the same consensus rules, which neglects all subjectivity that is
inherent to social content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illegal content (e.g., hate speech).

In this work, we adapt the idea of scarce resources to reach consensus, but in
the context of social content sharing.
%
Our first contribution is to recognize the actual published contents as the
protocol scarce resources, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are used to evaluate other posts with likes and
dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and made by multiple users.
%
Due to decentralization, posts in a timeline form a causal graph with only
partial order, which we promote to a total order based on the reputation of
authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow that users create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer content dissemination protocol that provides strong
eventual consistency~\cite{p2p.crdts,p2p.sec}.
%
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

@@

\newpage
\section{The Reputation and Consensus Mechanism}
\label{sec.design}

@@

We propose a reputation system in which users can spend tokens named \reps to
post and rate content in the forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the reputation operations and their goals.
To prevent Sybils, users with no \reps cannot perform these operations,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users approve,
which demands non-trivial work resistant to automation.%
\footnote{
    Note that to some extent work from AI may be indistinguishable from humans,
    but its frequency and relevance are still subject to human evaluation,
    which is subjective and evolves over time.
}

@@

Bitcoin employs proof-of-work to mitigate Sybil attacks.
However, CPU or other extrinsic resources are not evenly distributed among
humans, specially in communications using battery-powered devices.
%
Instead, considering the context of public forums, we can take advantage of the
human authoring ability as an intrinsic resource.
Creating new content is hard and takes time, but is comparatively easy to
verify and rate.
Therefore, in order to impose scarcity, we determine that only content
authoring generates \reps, while likes and dislikes just transfer \reps between
users.
%but limited by periods of time to prevent excess (e.g., once in a day per user).
%Additionally,
%
Nevertheless, scarce operations are not yet sufficient because they demand
consensus to establish an order in time across the network to prevent
inconsistent operations.
As an example, consider a malicious author with a single unit of \reps posting
new messages using multiple peers at the same time.
According to the \emph{Expense} rule of Table~\ref{fig.general}, only one of
these messages should be accepted.
However, without consensus, it is not possible to globally determine which
message to accept, since each peer would supposedly accept the first message it
sees.
On the one hand, in order to validate operations consistently, we need the same
message ordering across all peers in the network.
On the other hand, we typically can only represent peer-to-peer public forums
as DAGs with causal relationships between messages, which provides at most
partial order.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{reps2.png}
\caption{
    (A) A public forum DAG with a common prefix and two branches.
    (B) Total order between blocks of the DAG after consensus.
}
\label{fig.reps}
\end{figure}

Therefore, our key idea to stablish consensus in forum DAGs is to favor forks
with posts from users that constitute the majority of the reputation in the
network.
These forks have more associated work from active users and are analogous to
longest chains in Bitcoin.
%
In technical terms, we can adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the consensus criteria.
%, still abstractly, since we did not discuss the actual rules for content creation and rating.
A public forum DAG has a common prefix with signed posts from users $a$, $b$,
and $c$.
Let's assume that within the prefix, users $a$ and $b$ have contributed with
better content and have more reputation combined than $c$ has alone (i.e.,
$8+5>3$).
%
After the prefix, the forum forks in two branches:
in \code{branch-1}, only user $c$ remains active and we see that new users $x$
and $y$ (with no previous reputation in the common prefix) generate a lot of
new content;
in \code{branch-2}, only users $a$ and $b$ participate but with less activity.
Nevertheless, \code{branch-2} would be ordered first in this case because,
before the forking point, $a$ and $b$ have more reputation than $c$, $x$, and
$y$ combined.
%
User $c$ here might represent a malicious user trying to cultivate fake
identities $x$ and $y$ in separate of the network during weeks to accumulate
\reps.
%However, the whole malicious \code{branch-1} is vulnerable because users in
%\code{branch-2} with more previous reputation take the priority and can
%overthrow user $c$.

Figure~\ref{fig.reps}.B indicates the resulting consensus order between blocks
in the forum.
All operations in \code{branch-2} appear before any operation in
\code{branch-1}.
%
At any point in the consensus timeline, if an operation fails, all remaining
blocks in the offending branch are removed from the primary DAG.
As an example, suppose that the last post by $a$ (in gray) is a dislike to user
$c$.
Then, it's possible that the last post by $c$ (in red), now (suppose) with
\nreps{0}, is rejected together with all posts by $y$ and $x$ in sequence.
%
The removal rule for the remaining blocks is a direct consequence of tamper
proof Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} used in permissionless protocols,
which cannot support graph modifications.

Note that the consensus order exists only for accountability purposes, and is
a view of the primary DAG structure.
%
Note also that users in the branch with more reputation can react to attacks
even after the fact.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} so
that a further merge removes all blocks of \code{branch-1} from the DAG.
%
There are some other relevant considerations about forks and merges:
%
Peers that first received branches with less reputation will need to reorder
all blocks starting at the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when a peer
detects a new longest chain. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to put
the other branch in sequence with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of rejection when
rejoining.

As a counterpoint to the consensus order in Figure~\ref{fig.reps}.B, maybe
users $a$ and $b$ have abandoned the forum for months, and thus \code{branch-1}
is actually legit.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum.
Yet another possibility is that both branches are legit but became disconnected
for a long period of time.
In any case, it is unacceptable that a very old remote branch affects a long
active local forum.
%
For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, creating \emph{hard forks} in
the network.
A hard fork occurs when a local branch crosses a predetermined and irreversible
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches will
never synchronize again.
More than simply numeric disputes, hard forks represent social conflicts in
which reconciling branches is no longer possible.
%
Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
consensus}, which cannot be reordered, from \emph{unstable consensus}, which may
still be affected by incoming branches.
The activity threshold counts backwards, from the latest local block in the
unstable consensus.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{n100-d7.png}
\caption{
    Stable consensus freezes the order of blocks once they cross the threshold.
    The unstable order may still be affected by incoming branches.
}
\label{fig.hard}
\end{figure}

In summary, the rules to merge a branch from a remote machine \code{j} into a
branch from a local machine \code{i} are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix.
    \item otherwise, branches are ordered by an arbitrary criteria, such as
          lexicographical order of the block hashes immediately after the
          common prefix.
\end{itemize}

@@

Finally, we consider that revoking posts is fundamental in the context of
social content publishing, and thus, an important contribution of this work.
As described in Figure~\ref{fig.state}, a forum post has three possible states:
\code{BLOCKED}, \code{ACCEPTED}, or \code{REVOKED}.
%
If the author has reputation, a new post is immediately \code{ACCEPTED} in the
forum.
%
Otherwise, it is \code{BLOCKED} and requires a like from another user.
Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts and neither
retransmit them to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
A reasonable policy is to hold blocked posts in a temporary bag and retransmit
them for some visibility in the network.
%
Once accepted, a post becomes part of the forum and can never be removed
again, since Merkle~DAGs are immutable by design.
%Note that blocked posts that become accepted are always succeeded by a
%\emph{like} (Figure~\ref{fig.forum}).
%
However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
dislike than likes), the block becomes \code{REVOKED} and its payload is not
retransmitted to other peers.
Note that a block hash does not depend on its associated payload, but only on
the payload hash.
Hence, it is safe to remove the payload as long as one can prove its revoked
state.
Later, if the post receives new likes, it means that the payload is still known
somewhere and peers can request it when synchronizing again.

@@

Note that chains with the same name but different pioneers would be
incompatible because the hash of genesis blocks depend on the pioneers' public
keys.

@@

Evaluation item (a) measures the protocol overhead due to blockchain metadata,
which consists of a timestamp, author signature, and hashes (block id, payload,
backlinks, and likes).
%
The original chat archive is \texttt{800kB} in size, or \texttt{80B} for each
message, which includes a timestamp, a username, and the actual payload.
The simulated chat chain is \texttt{8MB} in size, which indicates a $10x$
overhead.
% (\code{20100825 [03:37:43]})
% (\code{hello everyone})
% (\code{<Ashlee>})
The original newsgroup is \texttt{30MB} in size, or \texttt{3kB} for each
message, which includes a timestamp, a sender, a subject, and the actual
payload (typically much longer).
The simulated newsgroup chain is \texttt{42MB} in size, which indicates a
$50\%$ overhead.
% (\code{29 Aug 88 23:36:58 GMT})
% (\code{mark@ramid.com})
% (\code{Error messages and Yacc})
%
It is clear that the metatada overhead is not negligible, specially for short
chat messages, but decreases as the payload increases.

Evaluation item (b) accounts the consensus algorithm applied locally at each
chain.
We measured the time to sort blocks in a local DAG both for the first time
(without any caches), and incrementally (from stable consensus caches).
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts respectively, while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts, regardless of the size
of the chain, which conveniently settles an upper bound on the input size of
the consensus algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

@@

Decentralized topic-based publish-subscribe protocols, such as
    \emph{XMPP}~\cite{pubsub.xmpp},
    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
peer-to-peer gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.stan,pubsub.vitis,pubsub.gossipsub,pubsub.rappel}.
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies blocks at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
In summary, \FC and \emph{pubsub} middlewares operate at different network
layers, which suggests that \FC could benefit of the latter to manage the
interconnections between peers.

>>>>

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in a consistent order, and
(iii) are respectful and on topic.
In a centralized system, items (i) and (ii) are trivially achieved assuming
availability and delivery order in the service, while for item (iii), users
have to trust the service to moderate content.
In a decentralized setting, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively replicate and
disseminate posts among peers until they reach all
users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish, \emph{at the protocol level}, between correct
and malicious users, and therefore, satisfy item (iii).
Without consensus, each client needs local anti-abuse policies, such as SPAM
filters and blacklists, but which only apply after the fact, when the abusive
messages have already been flooded in the network.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illegal content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybils through consensus.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our original decentralization goals, and more
importantly, they renounce subjectivity, which is inherent to content
moderation.
%
For instance, a unique timeline implies that all Internet content should be
subject to the same consensus rules, while objective rules alone cannot
distinguish abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illegal content (e.g., hate speech).
For instance, Bitcoin transactions permanently store arbitrary human-authored
content~\cite{btc.data}, including harmful texts and images~\cite{btc.content}.

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
%
Our first contribution is to recognize the actual published contents themselves
as the scarce resources, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are used to evaluate other posts with likes and
dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and made by multiple users.
%
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow users to create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer content dissemination protocol that provides strong
eventual consistency~\cite{p2p.crdts,p2p.sec}.
%
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

@@

\section{The Reputation and Consensus Mechanism}
\label{sec.design}

@@

@@

We propose a reputation system in which users can spend tokens named \reps to
post and rate content in the forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the reputation operations and their goals.
To prevent Sybils, users with no \reps cannot perform these operations,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users approve,
which demands non-trivial work resistant to automation.%
\footnote{
    Note that to some extent AI content may be indistinguishable from humans,
    but its frequency and relevance in forums are still subject to human
    evaluation in our proposal.
}

Bitcoin employs proof-of-work to mitigate Sybil attacks.
However, CPU or other extrinsic resources are not evenly distributed among
humans, specially in communications using battery-powered devices.
%
Instead, considering the context of public forums, we propose to take advantage
of the human authoring ability as an intrinsic resource.
Creating new content is hard and takes time, but is comparatively easy to
verify and rate.
Therefore, in order to impose scarcity, we determine that only content
authoring generates \reps, while likes and dislikes just transfer \reps between
users.
%but limited by periods of time to prevent excess (e.g., once in a day per user).
%Additionally,

Nevertheless, posts scarcity is not yet sufficient because they still demand a
consensual order in time across the network to prevent inconsistent operations.
As an example, consider a malicious author with a single unit of \reps posting
SPAM messages from multiple peers at the same time.
According to the \emph{Expense} rule of Table~\ref{fig.general}, only one of
these messages should be accepted by the network.
However, without consensus, it is not possible to globally determine which
message to accept, since each peer would supposedly accept the first message it
receives.
On the one hand, in order to validate operations consistently, we need the same
message ordering across all peers in the network.
On the other hand, due to network partitions, we can only represent public
forums as DAGs with causal relationships between messages, which provides at
most partial order.

\subsection{Basic Reputation Consensus}

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{reps.png}
\caption{
    (A) A public forum DAG with a common prefix, two branches, and a common sufix.
    (B) Total order between posts of the DAG after consensus.
}
\label{fig.reps}
\end{figure}

Our key idea to stablish consensus in forum DAGs is to favor forks with posts
from users that constitute the majority of the reputation in the network.
These forks have more associated work from active users and are analogous to
longest chains in Bitcoin.
%
In technical terms, we can adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
new users $x$ and $y$ with posts in \code{branch-1}.
In the meantime, $a$ and $b$ participate with posts in \code{branch-2}, but
with less activity.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches with a common suffix.

Even in the ordinary case of a legit temporary network partition, note that the
two branches may still show conflicts when merging.
For instance, posts with dislike operations in one branch could possibly block
posts from users in the other branch.
Therefore, we need to order the branches with a global consensus, such that we
can validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

In Figure~\ref{fig.reps}.A, we assume that within the common prefix, users $a$
and $b$ have contributed with better content and therefore have more reputation
combined than $c$ has alone (i.e., $8+5>3$).
%
Based on that information, Figure~\ref{fig.reps}.B indicates the resulting
consensus order between posts in the forum DAG:
    the common prefix comes first,
    followed by all posts in \code{branch-2},
    followed by all posts in \code{branch-1},
    followed by the common suffix.
%
The \code{branch-2} with users $a$ and $b$ take priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

At any point in the consensus timeline, if an operation fails, all remaining
posts in the offending branch are removed from the primary DAG.
As an example, suppose that the last post by $a$ (in gray) is a dislike to user
$c$.
Then, it's possible that the last post by $c$ (in red), now (suppose) with
\nreps{0}, is rejected together with all posts in sequence, including those in
the common suffix.
%
This rejection rule for remaining posts is a direct consequence of tamper proof
Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} typically used in permissionless
protocols, which cannot support graph modifications.

The basic consensus rule is therefore straightforward:
Whenever a fork is found, we create sets of users with posts in each branch
($B1=\{c,x,y\}$ and $B2=\{a,b\}$).
Then, we take each set, and sum its users \reps based on the common prefix
($S1=3$ and $S2=13$), since it is the maximum information that all branches
agree.
Finally, we order the branches from the highest sum of \reps found
(Figure~\ref{fig.reps}.B).
While applying the branches in order, if any post operation fails, all
remaining posts are rejected, as if they never existed.

There are some other relevant considerations about forks and merges:
%
Peers that first received branches with less reputation will need to reorder
all posts starting at the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when a peer
detects a new longest chain. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to put
the other branch in sequence with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on reputation of an agreed common prefix,
    (iv) leading to a global and deterministic and result.
%
Note that the consensus order exists only for accountability purposes, and is
a view of the DAG structure, which remains as the primary source of data.

\subsection{Malicious Behaviors and Hard Forks}

Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
intentionally disconnected from the network for weeks to cultivate fake
identities $x$ and $y$ and accumulate \reps.
When merging, these fake users become legit and now constitute the majority of
the forum in terms of reputation.
%
However, because the common prefix is the basis of consensus, users in the
branch with more reputation can still react even after the attack.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} before
the common suffix so that a further merge removes all posts of \code{branch-1}
from the DAG.
%
The example shows that this kind of attack is innocuous because it requires at
least half of the \reps accumulated in the common prefix.
%
The same countermeasure can be used against a minority of users colluding to
boost or protect their own posts.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{n100-d7.png}
\caption{
    Stable consensus permanently freezes the order of posts once they cross
    the threshold.
    The unstable order may still be affected by incoming branches.
}
\label{fig.hard}
\end{figure}

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legit.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum.
Yet another possibility is that both branches are legit but became disconnected
for a long period of time.
In any case, it is unacceptable for users that a very old remote branch affects
a long active forum.
%
For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, which leads to \emph{hard forks}
in the network.
A hard fork occurs when a local branch crosses a predetermined and irreversible
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches will
never synchronize again.
More than simple numeric disputes, a hard fork represents a social conflict in
which reconciling branches is no longer possible.
%
Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
consensus}, which cannot be reordered, from \emph{unstable consensus}, which
may still be affected by incoming branches.
The activity threshold counts backwards, starting from the newest local post
in the unstable consensus to older posts until they are permanently frozen.

In summary, the rules to merge a branch from a remote machine \code{j} into a
branch from a local machine \code{i} are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix.
    \item otherwise, branches are ordered by an arbitrary criteria, such as
          lexicographical order of the post hashes immediately after the
          common prefix.
\end{itemize}

@@

\subsection{Content Removal}

Finally, we consider that revoking posts is fundamental in the context of
social content publishing, and thus, an important contribution of this work.
As described in Figure~\ref{fig.state}, a forum post has three possible states:
\code{BLOCKED}, \code{ACCEPTED}, or \code{REVOKED}.
%
If the author has reputation, a new post is immediately \code{ACCEPTED} in the
forum.
%
Otherwise, it is \code{BLOCKED} and requires a like from another user.
Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts and neither
retransmit them to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
A reasonable policy is to hold blocked posts in a temporary bag and retransmit
them for some visibility in the network.
%
Once accepted, a post becomes part of the forum and can never be removed
again, since Merkle~DAGs are immutable by design.
%Note that blocked posts that become accepted are always succeeded by a
%\emph{like} (Figure~\ref{fig.forum}).
%
However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
dislike than likes), the post becomes \code{REVOKED} and its payload is not
retransmitted to other peers.
Note that a post hash does not depend on its associated payload, but only on
the payload hash.
Hence, it is safe to remove the payload as long as one can prove its revoked
state.
Later, if the post receives new likes, it means that the payload is still known
somewhere and peers can request it when synchronizing again.

\subsection{Byzantine Peers}

Since posts are signed and structured in Merkle~DAGs, ~\cite{p2p.ipfs,p2p.bitcoin} used in permissionless
Regarding Byzantine-faulty peers, which deviate from the protocol
p2p.byz
When a node deviates from the specified protocol, we
call it Byzantine-faulty (or simply Byzantine), regardless of
whether the deviation is by accident or by malice [25]. Byzan-
tine behaviour is not always detectable by other nodes, since
a Byzantine node may attempt to hide its protocol violations.

To conclude this section, we should keep in mind that our consensus algorithm
is based on subjective evaluations of users and posts.
For this reason, unlike Bitcoin, in which forks are discouraged and a single
blockchain must prevail, public forums can split and diverge indiscriminately.
For instance, users can explicitly apply hard forks at any point in time or
``reboot'' a community from a previous state.
Ultimately, the consensus algorithm provides a transparent mechanism to help
users understand the evolution of public forums and act accordingly.

@@

Note that chains with the same name but different pioneers would be
incompatible because the hash of genesis posts depend on the pioneers' public
keys.

@@

Evaluation item (a) measures the protocol overhead due to blockchain metadata,
which consists of a timestamp, author signature, and hashes (post id, payload,
backlinks, and likes).
%
The original chat archive is \texttt{800kB} in size, or \texttt{80B} for each
message, which includes a timestamp, a username, and the actual payload.
The simulated chat chain is \texttt{8MB} in size, which indicates a $10x$
overhead.
% (\code{20100825 [03:37:43]})
% (\code{hello everyone})
% (\code{<Ashlee>})
The original newsgroup is \texttt{30MB} in size, or \texttt{3kB} for each
message, which includes a timestamp, a sender, a subject, and the actual
payload (typically much longer).
The simulated newsgroup chain is \texttt{42MB} in size, which indicates a
$50\%$ overhead.
% (\code{29 Aug 88 23:36:58 GMT})
% (\code{mark@ramid.com})
% (\code{Error messages and Yacc})
%
It is clear that the metatada overhead is not negligible, specially for short
chat messages, but decreases as the payload increases.

Evaluation item (b) accounts the consensus algorithm applied locally at each
chain.
We measured the time to sort posts in a local DAG both for the first time
(without any caches), and incrementally (from stable consensus caches).
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts respectively, while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts, regardless of the size
of the chain, which conveniently settles an upper bound on the input size of
the consensus algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

@@

Decentralized topic-based publish-subscribe protocols, such as
    \emph{XMPP}~\cite{pubsub.xmpp},
    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
peer-to-peer gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.stan,pubsub.vitis,pubsub.gossipsub,pubsub.rappel}.
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies posts at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
In summary, \FC and \emph{pubsub} middlewares operate at different network
layers, which suggests that \FC could benefit of the latter to manage the
interconnections between peers.
